\bf Comparing the Power of Plot Designs to Reveal Correlation

### 2015-12-10 paper/paper.Rnw
getting closer to fixing latex errors
tweaking pin plot code

\documentclass[12pt]{article}
\usepackage{natbib}
\usepackage{color}
\usepackage[dvipsnames,svgnames*]{xcolor}
\usepackage{array}
\usepackage[colorlinks=TRUE, linkcolor=blue]{hyperref}
\usepackage{wrapfig,float}
\usepackage[font=small,skip=5pt]{caption}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{amsmath}
%\usepackage{dsfont}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{url}
\usepackage{ulem}
%\usepackage[section]{placeins}
\usepackage{afterpage}

@@

\newcommand{\blue}[1]{{\color{blue} #1}}
\newcommand{\hh}[1]{{\color{magenta} #1}}
\newcommand{\dc}[1]{{\color{orange} #1}}
\newcommand{\green}[1]{{\color{cyan} #1}}

@@

<<setup, echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
library(lme4)
library(ggplot2)
library(mnormt)
library(tidyr)
# Convert reshape2 to tidyr, melt to gather - done
#library(reshape2)
library(nullabor)
library(dplyr)
library(mFilter)
library(grid)
library(gridExtra)
#library(splitstackshape) - changed cSplit to unnest
#library(readr)
#library(lubridate)
# Convert these to read_csv at some point
correl <- read.csv("correlations.csv")
full <- read.csv("full.csv", row.names=1) 
Turk <- read.csv("turk.csv")

@@

\begin{figure}[htbp]
  \centering
  \includegraphics{cnyline}
  \caption{Exchange rates for Australian dollars and Chinese yuan against the United States dollar from Mar 30-May 2, 2015, shown as overlaid time series. (Both sets of values were standardized to bring them to the same scale for plotting.) What would you report the correlation between the two series to be?}
  \label{cny}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}
    \includegraphics{nzdscatter}
  \end{subfigure}
  \begin{subfigure}
    \includegraphics{nzdline}
  \end{subfigure}
  \caption{Exchange rates for Australian and NZ dollars against US dollar, from Mar 30-May 2, 2015, shown using a scatterplot (left) and overlaid line graph (right). Which one can you read the correlation from more accurately? Point-wise correlation between the two series, ignoring autodependence, is 0.64.}
  \label{nzd}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics{murders}
  \caption{[19] Murders in New York State (Orange) and New York City (Blue). The difference between the two represents murders outside of the city. What size would you judge the difference to be in 1994?}
  \label{murder}
\end{figure}

@@

If all $K$ observers pick the same plot out of a lineup of twenty, we result in a p-value as small as $1/m^K$ . This research used $m= 20$, giving us a significance level (and Type 1 error rate) of $α = 0.05$. The plot of 'true' data is placed randomly amongst 19 null plots to form the lineup.
\begin{figure}[htbp]
  \centering
<<lineup.example, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
set.seed(1234)
a <- gen_true_data(24, 0.95)
td <- gather(a, variable, value, c(X1, X2))
b <- gen_null(24)
b$.n <- NULL
nd_l <- gather(b, variable, value, c(X1, X2))
nd_l$.n <- rep(1:19, rep(24, 19))
pos <- 7
td <- data.frame(.n=rep(pos, nrow(td)), td)
lg <- nd_l$.n < pos
nd_l1 <- data.frame(.n=nd_l$.n[lg], t=nd_l$t[lg], variable=nd_l$variable[lg], 
                    value=nd_l$value[lg])
nd_l2 <- data.frame(.n=nd_l$.n[!lg], t=nd_l$t[!lg], variable=nd_l$variable[!lg], 
                    value=nd_l$value[!lg])
nd_l2$.n <- nd_l2$.n + 1
d <- rbind(nd_l1, td, nd_l2)

@@

\begin{figure}[htbp]
  \centering
<<smoothing, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
set.seed(123)
uns <- gen_true_data(96, 0.8)
uns_l <- gather(uns, variable, value, -t)
p1a <- ggplot(data=uns_l, aes(x=t, y=value, colour=variable)) + geom_line() +labs(title="Unsmoothed Overlaid Lines") +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p3a <- ggplot(data=uns, aes(x=X1, y=X2)) + geom_point()+ labs(title="Unsmoothed Scatterplot)") +
    theme(aspect.ratio=1, axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")

@@

\begin{figure}[htbp]
  \centering
<<designs, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
set.seed(101)
a <- gen_true_data(24, 0.95)
b <- gen_true_data(24, -0.95)
ca <- cor(a$X1, a$X2)
cb <- cor(b$X1, b$X2)
a_l <- gather(a, variable, value, -t)
b_l <- gather(b, variable, value, -t)
p1 <- ggplot(data=a_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
      theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p2 <- ggplot(data=b_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
      theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p3 <- ggplot(data=a, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
      theme(aspect.ratio=1, axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p4 <- ggplot(data=b, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
      theme(aspect.ratio=1, axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
grid.arrange(p1, p2, p3, p4, ncol=2)
@
  \label{designs}
  \caption{The same two data sets in both plot designs. t=24, correlation (left) = 0.946, correlation (right) = -0.943}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
%<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE>>=
\iffalse
# correl <- read.csv("correlations.csv") # You have already read this
pin <- select(full, pic_id, response_0, actual_location, 
              test_param, smoothed, n) #full[,c(5, 6, 10, 15, 16, 17)]
# Need to use the tidyr commands instead of cSplit
# Here's where we will need to make some changes to incorporate the multiple picks
pin <- pin %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
pin$response_0 <- as.integer(pin$response_0)
counts <- data.frame(table(pin$pic_id, pin$response_0))
counts$Var1 <- as.integer(counts$Var1)
counts$Var2 <- as.integer(counts$Var2)
colnames(counts) <- c("lineup", "plot", "count")
pin <- merge(counts, Turk, by.x="lineup", by.y="pic_id")
pin <- select(pin, lineup, plot, count, smoothed, true_pos, n, test_param, r)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
pin$r <- factor(pin$r)
colnames(pin) <- c("lineup", "plot", "count", "smoothed", "true_pos", "n", "design", "r")
pin <- mutate(pin, design=ifelse(design=="scatter", "Scatterplot", "Overlaid Lines"))#, smoothed=ifelse(smoothed=="no", 0, 1))
pin$base <- 0
pin$Correlation <- diag(as.matrix(correl[pin$lineup, (pin$plot+2)]))
pin <- mutate(pin, true = ifelse(plot==true_pos, "True plot", "Null plot"))
pin <- merge(pin, Turk[,c(2, 16)], by.x = "lineup", by.y = "pic_id")
pin <- mutate(pin, percent = count/attempt)
ggplot(data=subset(pin, smoothed=="no" & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
       geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed() + facet_grid(n~r) + labs(title="Unsmoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed=="yes" & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed=="no" & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Unsmoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed=="yes" & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")

#Code to be fixed
\fi
%@

@@

\begin{figure}[htbp]
  \centering
  \includegraphics{anom1}
  \label{anom1}
  \caption{A lineup of unsmoothed overlaid lines with a sample size of 12 and true correlation of -0.707. The true plot in position 8 was not detected in five evaluations. However, plot 13 was selected four times, (p –value < 0.001) despite having a correlation of only 0.225. When asked for a reason of their selections, subjects answered that the lines were closest together.}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics{anom2}
  \label{anom2}
  \caption{A lineup of unsmoothed scatter plots with a sample size of 24. The true plot is located in position 4 with a correlation of -0.491, and was selected by one out of eleven observers (p = 0.33, see formula \ref{eq:1}). The plot in position 6 was selected by nine of these eleven observers (p < 0.001) despite having a correlation of -0.037. This null plot shows a unique clustering effect that draws in the observer despite not having any linear relationship in the data, an example of the over-interpretation of randomness. Subjects most often cited the clustering pattern formed as the reason for their selection. The final selection was position 8, with a correlation of -0.226 (p = 0.33).}
\end{figure}

>>>>

\documentclass[12pt]{article}
\usepackage{natbib}
\usepackage{color}
\usepackage[dvipsnames,svgnames*]{xcolor}
\usepackage{array}
\usepackage[colorlinks=TRUE, linkcolor=blue]{hyperref}
\usepackage{wrapfig,float}
\usepackage[font=small,skip=5pt]{caption}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{dsfont}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{url}
\usepackage{ulem}
\usepackage[section]{placeins}
\usepackage{afterpage}

@@

@@

<<setup, echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
library(lme4)
library(ggplot2)
library(mnormt)
library(tidyr)
# Convert reshape2 to tidyr, melt to gather - done
#library(reshape2)
library(nullabor)
library(dplyr)
library(mFilter)
library(grid)
library(gridExtra)
#library(splitstackshape) - changed cSplit to unnest
#library(readr)
#library(lubridate)
# Convert these to read_csv at some point
correl <- read.csv("correlations.csv", stringsAsFactors = FALSE)
full <- read.csv("full.csv", row.names=1, stringsAsFactors = FALSE) 
Turk <- read.csv("turk.csv", stringsAsFactors = FALSE)

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\textwidth]{cnyline}
  \caption{Exchange rates for Australian dollars and Chinese yuan against the United States dollar from Mar 30-May 2, 2015, shown as overlaid time series. (Both sets of values were standardized to bring them to the same scale for plotting.) What would you report the correlation between the two series to be?}
  \label{cny}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
    \centerline{\includegraphics[width=0.45\textwidth]{nzdscatter}
    \includegraphics[width=0.45\textwidth]{nzdline}}
  \caption{Exchange rates for Australian and NZ dollars against US dollar, from Mar 30-May 2, 2015, shown using a scatterplot (left) and overlaid line graph (right). Which one can you read the correlation from more accurately? Point-wise correlation between the two series, ignoring autodependence, is 0.64.}
  \label{nzd}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\textwidth]{murders}
  \caption{[19] Murders in New York State (Orange) and New York City (Blue). The difference between the two represents murders outside of the city. What size would you judge the difference to be in 1994?}
  \label{murder}
\end{figure}

@@

If all $K$ observers pick the same plot out of a lineup of twenty, we result in a p-value as small as $1/m^K$ . This research used $m= 20$, giving us a significance level (and Type 1 error rate) of $α = 0.05$. The plot of 'true' data is placed randomly amongst 19 null plots to form the lineup.
\begin{figure}[htbp]
  \centering
<<lineup.example, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE>>=
set.seed(1234)
a <- gen_true_data(24, 0.95)
td <- gather(a, variable, value, c(X1, X2))
b <- gen_null(24)
b$.n <- NULL
nd_l <- gather(b, variable, value, c(X1, X2))
nd_l$.n <- rep(1:19, rep(24, 19))
pos <- 7
td <- data.frame(.n=rep(pos, nrow(td)), td)
lg <- nd_l$.n < pos
nd_l1 <- data.frame(.n=nd_l$.n[lg], t=nd_l$t[lg], variable=nd_l$variable[lg], 
                    value=nd_l$value[lg])
nd_l2 <- data.frame(.n=nd_l$.n[!lg], t=nd_l$t[!lg], variable=nd_l$variable[!lg], 
                    value=nd_l$value[!lg])
nd_l2$.n <- nd_l2$.n + 1
d <- rbind(nd_l1, td, nd_l2)

@@

\begin{figure}[htbp]
  \centering
<<smoothing, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE>>=
set.seed(123)
uns <- gen_true_data(96, 0.8)
uns_l <- gather(uns, variable, value, -t)
p1a <- ggplot(data=uns_l, aes(x=t, y=value, colour=variable)) + geom_line() +labs(title="Unsmoothed Overlaid Lines") +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p3a <- ggplot(data=uns, aes(x=X1, y=X2)) + geom_point()+ labs(title="Unsmoothed Scatterplot)") +
    theme(aspect.ratio=1, axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")

@@

\begin{figure}[htbp]
  \centering
<<designs, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE>>=
set.seed(101)
a <- gen_true_data(24, 0.95)
b <- gen_true_data(24, -0.95)
ca <- cor(a$X1, a$X2)
cb <- cor(b$X1, b$X2)
a_l <- gather(a, variable, value, -t)
b_l <- gather(b, variable, value, -t)
p1 <- ggplot(data=a_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
      theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p2 <- ggplot(data=b_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
      theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p3 <- ggplot(data=a, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
      theme(aspect.ratio=1, axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p4 <- ggplot(data=b, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
      theme(aspect.ratio=1, axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
grid.arrange(p1, p2, p3, p4, ncol=2)
@
  \label{designs}
  \caption{The same two data sets in both plot designs. t=24, correlation (left) = 0.946, correlation (right) = -0.943}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE>>=
#\iffalse
# correl <- read.csv("correlations.csv") # You have already read this
pin <- select(full, pic_id, response_0, actual_location, 
              test_param, smoothed, n) #full[,c(5, 6, 10, 15, 16, 17)]
# Need to use the tidyr commands instead of cSplit
# Here's where we will need to make some changes to incorporate the multiple picks
#pin <- pin %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
#pin$response_0 <- separate(pin, response_0, "response_0", sep=",", extra="drop")
pin$choice <- separate(pin, response_0, c("choice", "l1", "l2"), ",", extra="drop")$choice
pin$choice[str_detect(pin$actual_location, pin$response_0)] <- 
  pin$actual_location[str_detect(pin$actual_location, pin$response_0)]
                         #as.integer(pin$response_0)
counts <- data.frame(table(pin$pic_id, pin$choice))
counts$Var1 <- as.integer(counts$Var1)
counts$Var2 <- as.integer(counts$Var2)
colnames(counts) <- c("lineup", "plot", "count")
pin <- merge(counts, Turk, by.x="lineup", by.y="pic_id")
pin <- select(pin, lineup, plot.x, count, smoothed, true_pos, n, test_param, r)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
pin$r <- factor(pin$r)
colnames(pin) <- c("lineup", "plot", "count", "smoothed", "true_pos", "n", "design", "r")
pin <- mutate(pin, design=ifelse(design=="scatter", "Scatterplot", "Overlaid Lines")) #, smoothed=ifelse(smoothed=="no", 0, 1))
pin$base <- 0
pin$Correlation <- diag(as.matrix(correl[pin$lineup, (pin$plot+2)]))
pin <- mutate(pin, true = ifelse(plot==true_pos, "True plot", "Null plot"))
pin <- merge(pin, Turk[,c(2, 16)], by.x = "lineup", by.y = "pic_id")
pin <- mutate(pin, percent = count/attempt)
ggplot(data=filter(pin, smoothed==0 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
       geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed() + facet_grid(n~r) + labs(title="Unsmoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed==1 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed==0 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Unsmoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed==1 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
#Code to be fixed
#\fi
@

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\textwidth]{anom1}
  \label{anom1}
  \caption{A lineup of unsmoothed overlaid lines with a sample size of 12 and true correlation of -0.707. The true plot in position 8 was not detected in five evaluations. However, plot 13 was selected four times, (p –value < 0.001) despite having a correlation of only 0.225. When asked for a reason of their selections, subjects answered that the lines were closest together.}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\textwidth]{anom2}
  \label{anom2}
  \caption{A lineup of unsmoothed scatter plots with a sample size of 24. The true plot is located in position 4 with a correlation of -0.491, and was selected by one out of eleven observers (p = 0.33, see formula \ref{eq:1}). The plot in position 6 was selected by nine of these eleven observers (p < 0.001) despite having a correlation of -0.037. This null plot shows a unique clustering effect that draws in the observer despite not having any linear relationship in the data, an example of the over-interpretation of randomness. Subjects most often cited the clustering pattern formed as the reason for their selection. The final selection was position 8, with a correlation of -0.226 (p = 0.33).}
\end{figure}
